---
title: "GRLS Data Summary"
author: "Deanna Lanier"
date: "2024-04-12"
output: html_notebook
---

This notebook provides a data summary of the cleaned and categorized GRLS data from 2012 - 2020.

Note: Due to the quick turn around, Feedform, feedtype, Feedfreq, Feedamt, Feedamtunit, Feedlong, Feedlongunit were not included in the summary of the summary. Removed any entry with dogs weighing over 200.

After the removal of all duplicates and outlier entries there were a total of 16750 observations.

## Table 1: Summary of categorical data

There are alot of incomplete variables. Some of the variables with low completion percentages include bodyfat (30.5%), rawcook (56%), startdt (49.1%), stilluse (49.1), and enddt (49.1).

```{r}
character_datatable
```

## Table 2: Summary of numeric data

The numeric data also includes incomplete variables such as pct_homemade (4.7%) and pctdiet (48%). The table includes the mean and percentiles.

```{r}
Numeric_datatable
```

## Figure 1: Enrolled dogs

peak enrollment in 2014

```{r}
interactive_fig1
```

## Figure 2:BCS by year

Most common bcs is 5

```{r}
interactive_fig2
```

## Figure 3: Processing categorization

Most consume an ultraprocessed diet

```{r}
interactive_fig3
```

## Figure 4: Processing subcategorization

```{r}
interactive_fig4
```

## Figure 5: Frequency of processing subtypes and processing types

```{r}
interactive_fig5
```

## Figure 6: Processing diets by bcs

peak enrollment in 2014

```{r}
interactive_fig6
```


## Table 2: Key
**Table 1**

+--------------------------+-----------------+---------------------------------------------+
| **Category**             | **Subcategory** | **Definition**                              |
+==========================+=================+=============================================+
| Minimally Processed (MP) | -   CP-R        | -   Commercially Produced Raw               |
|                          | -   HM-R        | -   Homemade Raw                            |
|                          | -   CP-S        | -   Commercially Processed Supplemental Raw |
+--------------------------+-----------------+---------------------------------------------+
| Processed (P)            | -   CP-AD       | -   Commercially Processed Air Dried        |
|                          | -   CP-C        | -   Commercially Processed Cooled           |
|                          | -   CP-D        | -   Commercially Processed Dehydrated       |
|                          | -   CP-FD       | -   Commercially Processed Freeze Dried     |
|                          | -   HM-C        | -   Homemade Cooked                         |
+--------------------------+-----------------+---------------------------------------------+
| Ultra Processed (UP)     | -   CUP-D       | -   Commercially Ultra-Processed Dry        |
|                          | -   CUP-W       | -   Commercially Ultra-Processed Wet        |
+--------------------------+-----------------+---------------------------------------------+

: Table 1: Diet Processing Categorization and sub-categorizations. Diets are categorized by the level of processing. This is determined by evaluating the diet data provided by the veterinarians and dog owners.

## Figure 7: Top companies by year

```{r, out.width= "100%"}


final_plot
```

