---
title: "GRLS Data Summary - Figures"
author: "Deanna Lanier"
date: "2024-04-12"
output: html_document
---

This notebook provides a data summary of the cleaned and categorized GRLS data from 2012 - 2020.

Note: Due to the quick turn around, Feedform, feedtype, Feedfreq, Feedamt, Feedamtunit, Feedlong, Feedlongunit were not included in the summary of the summary. Removed any entry with dogs weighing over 200.

```{r}
#library 
library(dplyr) #for data processing/cleaning
library(ggplot2) #for figures
library(plotly) #for interactive figures
```

## Study population

### Figure 1 Study Enrollment

How many dogs were apart of the study each calendar year

```{r}
# Create a table that shows the frequency of the dogs enrolled by year

unique_count = Diet_Data_Clean %>%  
  distinct(calendaryear,id) %>%  #count unique values
  group_by(calendaryear)%>% # count the frequency by calendar year
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))# dont include any NA


```

Plot

```{r, warning=FALSE}
# plot the frequency 
figure1 = ggplot(data=unique_count, aes(x=calendaryear, y=Freq, group=1)) +
  geom_line()+
  geom_point()+ 
  theme_bw()+
  theme(axis.text.x = element_text(face="bold",  size=10)) + 
  xlab("Calendar Year") + 
  ylab("Number of Enrolled Goldens") +
  ggtitle("Study enrollment by year") +
  theme(
  plot.title = element_text(color="Black", size=14, face="bold"))+
  labs(caption = "Figure 1: Total study enrollment by year. The maximum number of participants was 3044 dogs in 2014.")+
  theme(plot.caption = element_text(hjust=0))
figure1

# Convert to an interactive Plotly plot
interactive_fig1 <- ggplotly(figure1)

# Print the interactive plot
interactive_fig1
```

### Figure 2 BCS by year

```{r}
#Create a table for the frequency of each Body condition score by year

BCS_table = Diet_Data_Clean %>%  
  distinct(calendaryear,bcs,id) %>% 
  group_by(calendaryear,bcs)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>% #no na
  filter(!is.na(bcs)) #no na
BCS_table
```

```{r}
#plot the frequency 

figure2 = ggplot(data=BCS_table, aes(x=calendaryear, y=Freq,  group=bcs, color=bcs)) +
  geom_line()+
  geom_point()+ 
  theme_bw()+
  theme(axis.text.x = element_text(face="bold",  size=10)) + 
  xlab("Year") + 
  ylab("Number of Dogs") +
  ggtitle("BCS by Year") +
  theme(
  plot.title = element_text(color="Black", size=14, face="bold"))+
  labs(caption = "Figure 2: Number of Goldens with specific body condition scores by year.")+
  theme(plot.caption = element_text(hjust=0))

figure2

# Convert to an interactive Plotly plot
interactive_fig2 <- ggplotly(figure2)

# Print the interactive plot
interactive_fig2
```

```{r}
#Create a table of the frequency of different diet types. Processing Categorization
Processing_table = Diet_Data_Clean %>%  
  distinct(calendaryear,process.category,id) %>%  
  group_by(calendaryear,process.category)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>%
  filter(!is.na(process.category))
Processing_table
```

```{r}
#plot that frequency 

figure3 = ggplot(data=Processing_table, aes(x=calendaryear, y=Freq,  group=process.category, color=process.category)) +
  geom_line()+
  geom_point()+ 
  theme_bw()+
  theme(axis.text.x = element_text(face="bold",  size=10)) + 
  xlab("Year") + 
  ylab("Number of Dogs") +
  ggtitle("Processing Categorization by Year") +
  theme(
  plot.title = element_text(color="Black", size=10, face="bold"),axis.text.x = element_text(angle = 45))+
  labs(caption = "Figure 3: Most goldens consumed an ultraprocessed diet for the duration of the study.")+
  theme(plot.caption = element_text(hjust=0))

figure3

# Convert to an interactive Plotly plot
interactive_fig3 <- ggplotly(figure3)

# Print the interactive plot
interactive_fig3
```

```{r}
# create frequency table of the processing subtypes 
ProcessingSub_table = Diet_Data_Clean %>%  
  distinct(calendaryear,process.subtype,id) %>%  
  group_by(calendaryear,process.subtype)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>%
  filter(!is.na(process.subtype))
ProcessingSub_table
```

```{r}
#Plot the subtypes 

figure4 = ggplot(data=ProcessingSub_table, aes(x=calendaryear, y=log(Freq),  group=process.subtype, color=process.subtype)) +
  geom_line()+
  geom_point()+ 
  theme_bw()+
  theme(axis.text.x = element_text(face="bold",  size=10)) + 
  xlab("Calendar Year") + 
  ylab("Number of Dogs (log)") +
  ggtitle("Processing Subtype Categorization by Year") +
  theme(
  plot.title = element_text(color="Black", size=10, face="bold"),axis.text.x = element_text(angle = 45))+
  labs(caption = "Figure 4: Most goldens consume CUP-D diet subtype.")+
  theme(plot.caption = element_text(hjust=0))


figure4

# Convert to an interactive Plotly plot
interactive_fig4 <- ggplotly(figure4)

# Print the interactive plot
interactive_fig4
```

```{r}
#create a frequency table of the processing subtypes grouped by process category and calendar year.
Processing_Sub_group = Diet_Data_Clean %>%
  distinct(calendaryear,process.category,process.subtype,id) %>%  
  group_by(calendaryear,process.category,process.subtype)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>%
  filter(!is.na(process.subtype))%>%
  filter(!is.na(process.category)) %>% 
  mutate(process.category = replace(process.category, process.category == "Minimally Processed", "MP")) %>% 
  mutate(process.category = replace(process.category, process.category == "Ultra Processed", "UP"))%>% 
  mutate(process.category = replace(process.category, process.category == "Processed", "P"))
Processing_Sub_group

```

```{r}
#Plot this by year 

figure5 = ggplot(Processing_Sub_group,                         # Draw barplot with grouping & stacking
       aes(x = process.category,
           y = log(Freq),
           fill = process.subtype)) + 
  geom_bar(stat = "identity",
           position = "stack") +
  facet_grid(~ calendaryear)+ 
  #theme(legend.position="bottom")+ 
  theme_bw()+
  theme(axis.text.x = element_text(size=8)) + 
  xlab("Processing Category") + 
  ylab("Number of Dogs (log)") +
  labs(fill = "Subcategory") +
  ggtitle("Frequency of Different Processed Subtypes by Processing Type") +
  theme(
  plot.title = element_text(color="Black", size=10, face="bold"),axis.text.x = element_text(angle = 45))+
  labs(caption = "Figure 4: Most goldens consume CUP-D diet subtype.")+
  theme(plot.caption = element_text(hjust=0))

figure5

# Convert to an interactive Plotly plot
interactive_fig5 <- ggplotly(figure5)

# Print the interactive plot
interactive_fig5
```

### Figure 5 BCS and processing categorization over the years

```{r}
#Create frequency table of the processing categorizations grouped by BCS and calendar year
BCS_Processingtable = Diet_Data_Clean %>%
  distinct(calendaryear,bcs,process.category,id) %>%  
  group_by(calendaryear,bcs,process.category)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>%
  filter(!is.na(bcs))%>%
  filter(!is.na(process.category)) %>% 
  mutate(process.category = replace(process.category, process.category == "Minimally Processed", "MP")) %>% 
  mutate(process.category = replace(process.category, process.category == "Ultra Processed", "UP"))%>% 
  mutate(process.category = replace(process.category, process.category == "Processed", "P"))
BCS_Processingtable
```

plot

```{r}
#create plot
figure6 = ggplot(BCS_Processingtable,                         # Draw barplot with grouping & stacking
       aes(x = bcs,
           y = log(Freq),
           fill = process.category)) + 
  geom_bar(stat = "identity",
           position = "stack") + theme_bw()+
  facet_grid(~ calendaryear)+ 
  theme(legend.position="bottom")+ 
  theme(axis.text.x = element_text(size=8)) + 
  xlab("BCS") + 
  ylab("Number of Dogs (log)") +
  labs(fill = "Processing") +
  ggtitle("Frequency of Different Processed Diets By BCS") +
  theme(
  plot.title = element_text(color="Black", size=10, face="bold"))+
  labs(caption = "Figure 5: Frequency of diets by established categorizations grouped by the body condition scores and year.")+
  theme(plot.caption = element_text(hjust=0))

figure6

# Convert to an interactive Plotly plot
interactive_fig6 <- ggplotly(figure6)

# Print the interactive plot
interactive_fig6
```

### Figure 7 Parent company

```{r}
#Create frequency table of the processing categorizations grouped by BCS and calendar year
Parentcompanytable = Diet_Data_Clean %>%
  distinct(calendaryear,Parent_Company,id) %>%  
  group_by(calendaryear,Parent_Company)%>% 
  summarize ("Freq" = n())%>%
  filter(!is.na(calendaryear))%>%
  filter(!is.na(Parent_Company)) 
Parentcompanytable

Parentcompany_datatable = datatable(Parentcompanytable, options = list(), class = "display")
Parentcompany_datatable
```

```{r}
# Filter to top 15 by Freq within each calendaryear
top_entries <- Parentcompanytable %>%
  group_by(calendaryear) %>%
  slice_max(order_by = Freq, n = 10) %>%
  ungroup()  # optional, to remove the grouping
top_entries
top_entries_datatable = datatable(top_entries, options = list(), class = "display")
top_entries_datatable
```

```{r}

# Prepare the data by selecting top 10 companies by frequency for each year
top_companies_by_year <- Parentcompanytable %>%
  group_by(calendaryear, Parent_Company) %>%
  summarise(Freq = sum(Freq), .groups = 'drop') %>%
  arrange(calendaryear, desc(Freq)) %>%
  group_by(calendaryear) %>%
  slice_max(order_by = Freq, n = 10) %>%
  ungroup() %>%
  arrange(calendaryear, desc(Freq))

top_companies_by_year

top_companies_by_year_datatable = datatable(top_companies_by_year, options = list(), class = "display")
top_companies_by_year_datatable
```

```{r}
# Create a list of plotly plots, one for each year
plots <- lapply(unique(top_companies_by_year$calendaryear), function(year) {
  filtered_data <- top_companies_by_year %>% 
    filter(calendaryear == year)
  
  p <- plot_ly(data = filtered_data, x = ~Parent_Company, y = ~Freq, type = 'bar',
               name = year, text = ~Freq, textposition = 'auto') %>%
    layout(title = paste("Top 10 Parent Companies in", year),
           xaxis = list(title = "Parent Company"),
           yaxis = list(title = "Frequency"))
  
  return(p)
})

# Create a plot with buttons to switch between years
final_plot <- subplot(plots, nrows = 1, shareY = TRUE) %>%
  layout(
    updatemenus = list(
      list(
        type = "buttons",
        direction = "left",
        buttons = lapply(1:length(unique(top_companies_by_year$calendaryear)), function(i) {
          list(method = "restyle",
               args = list("visible", as.logical(c(rep(FALSE, length(plots)), TRUE)[(i-1)*length(plots) + (1:length(plots))])),
               label = unique(top_companies_by_year$calendaryear)[i])
        }),
        pad = list(r = 10, t = 10),
        showactive = TRUE,
        x = 0.10,
        xanchor = "left",
        y = 1.1,
        yanchor = "top"
      )
    )
  )
```

```{r}
final_plot
```

```{r}

test <- top_companies_by_year %>% 
    filter(calendaryear == year) %>%
    arrange(desc(Freq)) %>%
    slice_max(n = 10, order_by = Freq)

# Create a list of plotly plots, one for each year
plots <- lapply(unique(top_companies_by_year$calendaryear), function(year) {
  filtered_data <- top_companies_by_year %>% 
    filter(calendaryear == year) %>%
    arrange(desc(Freq)) %>%
    slice_max(n = 10, order_by = Freq) # Ensure top 10 are selected, sorted if needed

  p <- plot_ly(data = filtered_data, x = ~Parent_Company, y = ~Freq, type = 'bar',
               name = year, text = ~Freq) %>%
    layout(title = paste("Top 10 Parent Companies in", year),
           xaxis = list(title = "Parent Company", tickangle = -45, tickfont = list(size = 8)),
           yaxis = list(title = "Frequency"))
  
  return(p)
})

# Create a plot with buttons to switch between years
final_plot <- subplot(plots, nrows = 1, shareY = TRUE) %>%
  layout(
    updatemenus = list(
      list(
        type = "buttons",
        direction = "left",
        buttons = lapply(1:length(unique(top_companies_by_year$calendaryear)), function(i) {
          list(method = "restyle",
               args = list("visible", as.logical(c(rep(FALSE, length(plots) * (i-1)), TRUE, rep(FALSE, length(plots) * (length(plots) - i))))),
               label = unique(top_companies_by_year$calendaryear)[i])
        }),
        pad = list(r = 10, t = 10),
        showactive = TRUE,
        x = 0.10,
        xanchor = "left",
        y = 1.1,
        yanchor = "top"
      )
    )
  )

# Print final_plot to display it
final_plot
```

```{r}
save(interactive_fig1,interactive_fig2,interactive_fig3,interactive_fig4,interactive_fig5,interactive_fig6,final_plot, file = "figures.RData")

save(character_datatable,Numeric_datatable, file = "tables.RData")

```

