---
title: "Data Selection 2"
author: "Deanna Lanier"
date: "2024-04-20"
output: html_document
---

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(dplyr)
```

```{r, echo = FALSE}
project.path = here()
#Load 
load("~/Dropbox (Edison_Lab@UGA)/Projects/vet/Deanna/Golden_Data/figures4.RData")
load("~/Dropbox (Edison_Lab@UGA)/Projects/vet/Deanna/Golden_Data/tables4.RData")

```

### 1. Table 1 & Figure 1 Unique Dogs by study Year

How many individual dogs were in each study year (not calendar year)?

```{r}
unique_count_datatable
```

```{r, echo = FALSE}
interactive_fig1
```

### 2. Table 2 & Figure 2 : BCS ranges by study year

• Per study year: how many of those individual dogs had a BCS of 0-3

• Per study year: how many of those individual dogs had a BCS of 4-6

• Per study year: how many of those individual dogs had a BCS of 7-9

```{r, echo = FALSE}
BCS_datatable
```

```{r, echo = FALSE}
interactive_fig2
```

### Table 3 & Figure 3: Processing Categorization by study year

```{r, echo = FALSE}
processing_datatable
```

```{r, echo = FALSE}
interactive_fig3
```

## Disease states

### Major Diagnosis over Study years

Cardiovascular Endocrine Gastrointestinal Hematologic Musculoskeletal Nervous Urinary

#### Fig 4 - grouped by BCS

```{r, echo = FALSE}
interactive_fig4
```

#### Table - Frequency of diagnosed v not diagnosed by study year

```{r, echo = FALSE}
major_diagnosis_summary_datatable

```

### Minor Diagnosis over Study years

Eye Ear nose throat infectious skin dental reproductive

#### Fig 4 - grouped by BCS

```{r, echo = FALSE}
interactive_fig5
```

#### Table - Frequency of diagnosed v not diagnosed by study year

```{r, echo = FALSE}
minor_diagnosis_summary_datatable
```

### All diagnosis over Study years (excluding dental and reproductive)

Cardiovascular Endocrine Gastrointestinal Hematologic Musculoskeletal Nervous Urinary Eye Ear nose throat infectious skin

#### Fig 4 - grouped by BCS

```{r, echo = FALSE}
interactive_fig6
```

#### Table - Frequency of diagnosed v not diagnosed by study year

```{r, echo = FALSE}
diagnosis_summary_datatable
```

### Eligible dogs based on age, bcs, and diagnosis

#### 49 dogs with the following criteria

There are *49* dogs that meet the following criteria:

1.  Minimally Processed Diet
2.  BCS = 4-6 3
3.  No major diagnosis (generally healthy)
4.  Entries across age groups (0-3, 4-6, and 7+)

```{r, echo = FALSE}
eligible_dogs1_datatable
```


#### 1078 Dogs with the following

There are *1078* dogs that meet the following criteria:

1.  Ultra Processed Diet
2.  BCS = 4-6 
3.  No major diagnosis (generally healthy)
4.  Entries across age groups (0-3, 4-6, and 7+)

\*excel

#### 1561 dogs with the following

there are *1561* dogs that meet the following criteria: 

1.  Ultra Processed Diet
2.  BCS = 4-6 
3.  Entries across age groups (0-3, 4-6, and 7+)

\*excel

#### 96 dogs with the following

there are *96* dogs that meet the following criterial: 1. ultra processed diet 2. BCS 7+ 3. Maintained ultra processed diet from ages 0-3, 4-6, and 7+

1.  Ultra Processed Diet
2.  BCS = 7+
3.  Entries across age groups (0-3, 4-6, and 7+)

```{r, echo = FALSE}
eligible_dogs3_datatable
```

#### 213 dogs with the following

1.  ultra processed diet
2.  BCS 4-6 at age 0-3
3.  BCS 7+ at age 4-6 and 7+ Maintained ultra processed diets from ages 0-3, 4-6, and 7+

```{r, echo = FALSE}
eligible_dogs4_datatable
```
