---
title: "Sample Selection"
output: html_notebook
---

To answer:

*Comparing Young and old*
How many dogs 0-2 have no diagnosis? and do not have a bcs above 7
How many dogs 3-6 have no diagnosis? and do not have a bcs above 7
How many dogs 7+ have no diagnosis? and do not have a bcs above 7

*independent of diagnosis:*
How many dogs were BCS 8-9 at have 0-3 that also have an entry at age 7+? Are there any diagnosis?
Is there a suggicient number of dogs for a control group (ideally no diagnosis) of healthy-weight dogs (BCS 4-6)

*Independent of diagnosis and BCS:*
How many dogs were weating (mostly) minimally processed foods at ages 0-3?
How many dogs were eating (mostly) processed foods at ages 0-3?
How many of those same dogs eat that diet at ages 5 or 6+?

*Hemangiosarcoma and lymphoma:*

Neuter status and sex for both disease states


## Load Packagse

```{r}
library(readxl) #for loading Excel files

library(dplyr) #for data processing/cleaning

library(tidyr) #for data processing/cleaning

library(skimr) #for nice visualization of data 

library(here) #to set paths

library(DT) #nice datatables 

```

## Load data
```{r}
# Restore the object
Diet_Data_Clean = readRDS(file = "~/Dropbox (Edison_Lab@UGA)/Projects/vet/Deanna/Golden_Data/CleanDietData.rds")
Death_Data_Clean = readRDS(file = "~/Dropbox (Edison_Lab@UGA)/Projects/vet/Deanna/Golden_Data/CleanDeathData.rds")
```


## Comparying Young v old

*How many dogs 0-2 have no diagnosis? and do not have a bcs above 7*
```{r}
# Convert 'BCS' to numeric
Diet_Data_Clean$BCS_numeric <- as.numeric(as.character(Diet_Data_Clean$bcs))
```

```{r}
# Filter Diet_Data_Clean for age and BCS conditions
#eligible_young_dogs <- Diet_Data_Clean %>%
#  filter(visit.age < 3, BCS_numeric < 7)
#length(unique(eligible_young_dogs$id))

# Filter Diet_Data_Clean for age under 3 years and BCS below 7
eligible_young_dogs <- Diet_Data_Clean %>%
  filter(visit.age < 3, BCS_numeric < 7) %>%
  select(id, study_year) %>%
  distinct()
length(unique(eligible_young_dogs$id))

```

```{r}
# Create a vector of ids from Death_Data_Clean that match the study year
#diagnosed_ids <- Death_Data_Clean %>%
#  select(id, study_year) %>%
#  distinct() %>%
#  pull(id)  # This creates a vector of unique ids

# Create a data frame of ids and study_years from Death_Data_Clean
diagnosed_data <- Death_Data_Clean %>%
  select(id, study_year) %>%
  distinct()

```

```{r}
# Remove dogs that have been diagnosed from the eligible dogs
#final_young_dogs <- eligible_young_dogs %>%
#  filter(!(id %in% diagnosed_ids))

# Identify dogs from eligible_young_dogs that have not been diagnosed in the same study year
final_young_dogs <- eligible_young_dogs %>%
  anti_join(diagnosed_data, by = c("id", "study_year"))

```

```{r}
length(unique(final_young_dogs$id))
```

Check diagnosis data to be sure
```{r}
death_datatable = datatable(Death_Data_Clean, options = list(), class = "display")
#death_datatable
```


*How many dogs 3-6 have no diagnosis? and do not have a bcs above 7*



```{r}
# Filter Diet_Data_Clean for age and BCS conditions
#eligible_middle_dogs <- Diet_Data_Clean %>%
#  filter(visit.age >= 3, visit.age <7, BCS_numeric < 7)
#length(unique(eligible_middle_dogs$id))

# Filter Diet_Data_Clean for age under 3 years and BCS below 7
eligible_middle_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 3, visit.age <7, BCS_numeric < 7) %>%
  select(id, study_year) %>%
  distinct()
length(unique(eligible_middle_dogs$id))

```

```{r}
# Create a vector of ids from Death_Data_Clean that match the study year
#diagnosed_ids <- Death_Data_Clean %>%
#  select(id, study_year) %>%
 # distinct() %>%
#  pull(id)  # This creates a vector of unique ids

# Create a data frame of ids and study_years from Death_Data_Clean
diagnosed_data <- Death_Data_Clean %>%
  select(id, study_year) %>%
  distinct()
```

```{r}
# Remove dogs that have been diagnosed from the eligible dogs
#final_middle_dogs <- eligible_middle_dogs %>%
#  filter(!(id %in% diagnosed_ids))

# Identify dogs from eligible_young_dogs that have not been diagnosed in the same study year
final_middle_dogs <- eligible_middle_dogs %>%
  anti_join(diagnosed_data, by = c("id", "study_year"))

```

```{r}
length(unique(final_middle_dogs$id))
```

Check diagnosis data to be sure
```{r}
death_datatable = datatable(Death_Data_Clean, options = list(), class = "display")
#death_datatable
```


*How many dogs 7+ have no diagnosis? and do not have a bcs above 7*

```{r}
# Filter Diet_Data_Clean for age and BCS conditions
#eligible_old_dogs <- Diet_Data_Clean %>%
#  filter(visit.age >7, BCS_numeric < 7)
#length(unique(eligible_old_dogs$id))

# Filter Diet_Data_Clean for age under 3 years and BCS below 7
eligible_old_dogs <- Diet_Data_Clean %>%
  filter(visit.age >7, BCS_numeric < 7) %>%
  select(id, study_year) %>%
  distinct()
length(unique(eligible_old_dogs$id))


```

```{r}
# Create a vector of ids from Death_Data_Clean that match the study year
diagnosed_ids <- Death_Data_Clean %>%
  select(id, study_year) %>%
  distinct() %>%
  pull(id)  # This creates a vector of unique ids

# Create a data frame of ids and study_years from Death_Data_Clean
#diagnosed_data <- Death_Data_Clean %>%
#  select(id, study_year) %>%
#  distinct()
```

```{r}
# Remove dogs that have been diagnosed from the eligible dogs
final_old_dogs <- eligible_old_dogs %>%
  filter(!(id %in% diagnosed_ids))

# Identify dogs from eligible_young_dogs that have not been diagnosed in the same study year
#final_old_dogs <- eligible_old_dogs %>%
#  anti_join(diagnosed_data, by = c("id", "study_year"))


# Count the number of unique dogs in the final_old_dogs data frame
final_count <- length(unique(final_old_dogs$id))

# Print the final count
print(final_count)

```

```{r}
length(unique(final_old_dogs$id))
```

Check diagnosis data to be sure
```{r}
death_datatable = datatable(Death_Data_Clean, options = list(), class = "display")
death_datatable
```


*How many dogs have datapoints in all 3 tables 
```{r}
# Extract unique IDs from each dataset
young_ids <- distinct(final_young_dogs, id) %>% pull(id)
middle_ids <- distinct(final_middle_dogs, id) %>% pull(id)
old_ids <- distinct(final_old_dogs, id) %>% pull(id)

# Find common IDs in all three datasets
common_ids <- Reduce(intersect, list(young_ids, middle_ids, old_ids))

# Count the number of unique dogs present in all three tables
common_count <- length(common_ids)

# Print the number of dogs
print(common_count)
```


## independent of diagnosis:

*How many dogs were BCS 8-9 at age 0-3 that also have an entry at age 7+? Are there any diagnosis?*

```{r}
#Filter dogs with BCS 8-9 at ages 0-3
young_bcs_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 0 & visit.age <= 3, bcs %in% c(8, 9)) %>%
  distinct(id)  # Assuming BCS is numeric

# Find those dogs with records at age 7+
older_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 7, id %in% young_bcs_dogs$id) %>%
  distinct(id)

#diagnosed_data <- Death_Data_Clean %>%
#  select(id, study_year) %>%
#  distinct()

# Create a vector of ids from Death_Data_Clean that match the study year
diagnosed_ids <- Death_Data_Clean %>%
  select(id, study_year) %>%
  distinct() %>%
  pull(id)  # This creates a vector of unique ids

# Remove dogs that have been diagnosed from the eligible dogs
diagnosed_dogs <- older_dogs %>%
  filter((id %in% diagnosed_ids))


#anti_join(diagnosed_data, by = c("id", "study_year"))


# Count the number of diagnosed dogs from the older_dogs group
diagnosed_count <- nrow(diagnosed_dogs)

# Print the count of diagnosed dogs
print(diagnosed_count)
```

Is there a suggicient number of dogs for a control group (ideally no diagnosis) of healthy-weight dogs (BCS 4-6)

```{r}
#Filter dogs with BCS 8-9 at ages 0-3
young_control_bcs_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 0 & visit.age <= 3, bcs %in% c(4,5,6)) %>%
  distinct(id)  # Assuming BCS is numeric

# Find those dogs with records at age 7+
olde_controlr_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 7, id %in% young_control_bcs_dogs$id) %>%
  distinct(id)

#diagnosed_data <- Death_Data_Clean %>%
#  select(id, study_year) %>%
#  distinct()

# Create a vector of ids from Death_Data_Clean that match the study year
diagnosed_ids <- Death_Data_Clean %>%
  select(id, study_year) %>%
  distinct() %>%
  pull(id)  # This creates a vector of unique ids

# Remove dogs that have been diagnosed from the eligible dogs
diagnosed_dogs <- olde_controlr_dogs %>%
  filter(!(id %in% diagnosed_ids))


#anti_join(diagnosed_data, by = c("id", "study_year"))


# Count the number of diagnosed dogs from the older_dogs group
diagnosed_count <- nrow(diagnosed_dogs)

# Print the count of diagnosed dogs
print(diagnosed_count)
```


##Independent of BCS/Diagnoses:

How many dogs were eating (mostly) minimally processed foods ages 0-3? Add processed if we have to if there are not enough MP. Just let me know how many were MP v P.

```{r}
# Filter the Diet_Clean_Data for dogs aged 0-3 and having a mostly minimally processed diet
minimally_processed_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 0 & visit.age <= 3, process.category == "Minimally Processed") %>%
  distinct(id)  # Assuming each dog's id is unique per entry

# Count the number of unique dogs that fit this criteria
count_minimally_processed_dogs <- nrow(minimally_processed_dogs)

# Print the number of dogs
print(count_minimally_processed_dogs)
```


How many of those same dogs were still eating (mostly) minimally processed Age 5 or 6+?

```{r}
# Filter the Diet_Clean_Data for dogs aged 0-3 and having a mostly minimally processed diet
minimally_processed_older_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 5 & visit.age <= 6, process.category == "Minimally Processed") %>%
  distinct(id)  # Assuming each dog's id is unique per entry

# Count the number of unique dogs that fit this criteria
count_minimally_processed_older_dogs <- nrow(minimally_processed_older_dogs)

# Print the number of dogs
print(count_minimally_processed_older_dogs)
```

How many dogs were eating (mostly) processed foods ages 0-3? Add processed if we have to if there are not enough MP. Just let me know how many were MP v P.

```{r}
# Filter the Diet_Clean_Data for dogs aged 0-3 and having a mostly minimally processed diet
processed_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 0 & visit.age <= 3, process.category == "Processed") %>%
  distinct(id)  # Assuming each dog's id is unique per entry

# Count the number of unique dogs that fit this criteria
count_processed_dogs <- nrow(processed_dogs)

# Print the number of dogs
print(count_processed_dogs)
```


How many of those same dogs were still eating (mostly)  processed Age 5 or 6+?

```{r}
# Filter the Diet_Clean_Data for dogs aged 0-3 and having a mostly minimally processed diet
processed_older_dogs <- Diet_Data_Clean %>%
  filter(visit.age >= 5 & visit.age <= 6, process.category == "Processed") %>%
  distinct(id)  # Assuming each dog's id is unique per entry

# Count the number of unique dogs that fit this criteria
count_processed_older_dogs <- nrow(processed_older_dogs)

# Print the number of dogs
print(count_processed_older_dogs)
```

```{r}
#Identify dogs that ate minimally processed diets at both age ranges
dogs_both_age_ranges <- intersect(processed_older_dogs$id, processed_dogs$id)

# Count the number of unique dogs that meet both criteria
count_dogs_both_age_ranges <- length(dogs_both_age_ranges)
count_dogs_both_age_ranges
```

## hemangiosarcoma and lymphoma

neuter status
```{r}
# Extract IDs and study years for dogs diagnosed with hemangiosarcoma or lymphoma
diagnosed_dogs <- Death_Data_Clean %>%
  filter(condition %in% c("hemangiosarcoma", "lymphoma")) %>%
  select(id, study_year) %>%
  distinct()

# Join with Diet_Data_Clean to get neuter status and sex
dog_details <- diagnosed_dogs %>%
  left_join(Diet_Data_Clean, by = c("id", "study_year")) %>%
  select(id, study_year, intact.y.n, sex)

# Count neutered/spayed vs. not
neuter_status_counts <- dog_details %>%
  filter(!is.na(intact.y.n)) %>%  # Ensuring that data exists for neuter status
  count(intact.y.n)

neuter_status_counts
```

sex
```{r}

```

